---
title: "VNSO Trade Dashboard"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
runtime: shiny
resource_files: www/logo.png
---

```{r setup, include=FALSE, warning=FALSE}
library(knitr)
library(flexdashboard)
library(DBI)
library(ggplot2)
library(gganimate)
library(plotly)
library(shiny)
library(dplyr)
library(scales)
library(tidyverse)
library(gifski)
library(openxlsx)
library(DT)
library(countrycode) # Get country codes
library(rmapshaper) # Simplifying polygons
library(leaflet) # Interactive mapping
library(tidyr) #Table Pivoting
library(spatstat)
library(RSQLite)
library(DBI)

```

```{r}
# Note where VNSO code/data is on current computer
repository <- file.path(dirname(rstudioapi::getSourceEditorContext()$path), "..")
#setwd(repository) # Required for file.choose() function
```


```{r}
# Note the secure data path
secureDataFolder <- file.path(repository, "data", "secure")
secureDataFolder_sqlite <- file.path(repository, "data", "secure")
```

```{r}
# Note the open data path
openDataFolder <- file.path(repository, "data", "open")


#mydb <- dbConnect(RSQLite::SQLite(), "../data/secure/my-db.sqlite")

mydb <- dbConnect(RSQLite::SQLite(), file.path(secureDataFolder_sqlite, "my-db.sqlite"))

```

Dashboard
=======================================================================


Column {data-width=650}
-----------------------------------------------------------------------

```{r View Export Summary}

export <- dbGetQuery(mydb, "SELECT sum([Stat..Value])/1000000 as Totex FROM curtrade WHERE CP4=1000 or CP4=1021 or CP4=1022 or CP4=3071")

import <- dbGetQuery(mydb, "SELECT sum([Stat..Value])/1000000 as Totim FROM curtrade WHERE CP4=4000 or CP4=4071 or CP4=7100")

curtradebalance <- export - import

```

column {data-width=400}
------------------------------------------------------------------
### Total Current Month Export (Millions)

```{r Current trade moments}

valueBox((export),
         icon = 'fa-balance-scale',
         color = "green")
```

### Total Current Month Import (Millions)
```{r Current trade imports}
valueBox((import),
         icon = 'fa-balance-scale',
         color = "blue")

```

### Total Current Month Trade Balance (Millions)
```{r Current Month trade balance}
valueBox((curtradebalance),
         icon = 'fa-balance-scale',
         color = "red")

```


column {data-width=400}
------------------------------------------------------------------

```{r Trade in services}
services <- dbGetQuery(mydb, "SELECT * FROM tblservices ORDER BY Year DESC")
services_select <- services %>%
                      head(1)

exportServices <- services_select$export
importServices <- services_select$import
serviceBalance <- services_select$balance

```

### Services Export - 2019
```{r}
valueBox(exportServices,
         icon = 'fa-balance-scale',
         color = "green")
```


### Services Import - 2019
```{r}

valueBox(importServices,
         icon = 'fa-balance-scale',
         color = "blue")
```


### Services Trade Balance - 2019
```{r}

valueBox(serviceBalance,
         icon = 'fa-balance-scale',
         color = "green")
```


Column {data-width=400}
-----------------------------------------------------------------------

### Trades in Goods and Services by Year (Billions)

```{r Trade in goods and serviuces by year}
#Extract Trade in Goods from the IMTS records
tradesImport <- dbGetQuery(mydb, "SELECT Year, round(sum(Value)/1000000) as goodsImports FROM importHistory GROUP BY Year ORDER BY Year")
tradesExport <- dbGetQuery(mydb, "SELECT Year, round(sum(Value)/1000000) as goodsExports FROM exportHistory GROUP BY Year ORDER BY Year")

#Merge trade in goods of Import and Export
tradeGoods <- merge(tradesImport, tradesExport, by = "Year")
tradeGoods$goodsBalance <- tradeGoods$goodsExports - tradeGoods$goodsImports

#Exract Trade in Services from RBV table record
tradeServices <- dbGetQuery(mydb, "SELECT Year, importServices, exportServices, servicesBalance as servicesBalance FROM tblservices GROUP BY Year ORDER BY Year")

#Merge both trade in Services and Trade in Goods
tradeGoodsServices <- merge(tradeGoods, tradeServices, by = "Year")
tradeGoodsServices$overallBalance <- tradeGoodsServices$goodsBalance + tradeGoodsServices$servicesBalance


```

```{r goods exports/imports/balance by year interactive plot}
tradeGoodsServices %>%
  
  plot_ly(x = ~Year,
          y = ~goodsExports,
          color = "blue",
          type = "scatter",
          mode = "lines+markers",
          name = "Export of Goods") %>%
  
  add_trace(x = ~Year,
            y = ~goodsImports,
            color = "red",
            name = "Import of Goods") %>%
  
  add_trace(x = ~Year,
            y = ~exportServices,
            color = "green",
            name = "Export of Services") %>%
  
  add_trace(x = ~Year,
            y = ~importServices,
            color = "yellow",
            name = "Import of Services") %>%
  
  plotly::layout(xaxis = list(title = "Year"), yaxis = list(title = 'Value (Millions(Vatu))'))
```


### Trade Balances by Year (Millions)

```{r services exports/imports/balance by year interactive plot}

tradeGoodsServices %>%
  
  plot_ly(x = ~Year,
          y = ~goodsBalance,
          color = "blue",
          type = "scatter", 
          mode = "lines+markers",
          name = "Goods Balance") %>%
  
  add_trace(x = ~Year,
            y = ~servicesBalance,
            color = "red",
            name = "Services Balance") %>%
  
  add_trace(x = ~Year,
            y = ~overallBalance,
            color = "green",
            name = "Balance of External Goods & Services") %>%
  
  plotly::layout(xaxis = list(title = "Year"), yaxis = list(title = 'Value (Billion(Vatu))'))
```

Column
---------------------------------------------------------------------
### Trade in Goods by Ports

```{r Trades Ports in Billions}
#Ports Export computation
portsExport <- dbGetQuery(mydb, "SELECT Year, Office, Procedure, Value/1000000 as exportValue, 
                          CASE 
                              WHEN Office = 'VAIR' THEN 'VilaGoodsExport'
                              WHEN Office = 'VAIRP' THEN 'VilaGoodsExport'
                              WHEN Office = 'VILA' THEN 'VilaGoodsExport'
                              WHEN Office = 'VPOST' THEN 'VilaGoodsExport'
                              WHEN Office = 'VSEA' THEN 'VilaGoodsExport'
                              WHEN Office = 'VSEAP' THEN 'VilaGoodsExport'
                          ELSE
                              'SantoGoodsExport'
                          
                          END AS tradeLocation
                          FROM exportHistory
                          WHERE Procedure = '1000' or Procedure = '1021' or Procedure = '1022' or Procedure                                           = '3071' ")
dbWriteTable(mydb, "portsExport", portsExport, overwrite=TRUE)

portsExport_new <- dbGetQuery(mydb, "SELECT tradeLocation, Year, sum(exportValue) as exportValue FROM portsExport GROUP BY tradeLocation, Year ORDER BY Year ")

portsExport_new_pivot <- portsExport_new %>%
                         pivot_wider(names_from = tradeLocation, values_from = exportValue)


#Ports Import computation
portsImport <- dbGetQuery(mydb, "SELECT Year, Office, Procedure, Value/1000000 as importValue, 
                          CASE 
                              WHEN Office = 'VAIR' THEN 'VilaGoodsImport'
                              WHEN Office = 'VAIRP' THEN 'VilaGoodsImport'
                              WHEN Office = 'VILA' THEN 'VilaGoodsImport'
                              WHEN Office = 'VPOST' THEN 'VilaGoodsImport'
                              WHEN Office = 'VSEA' THEN 'VilaGoodsImport'
                              WHEN Office = 'VSEAP' THEN 'VilaGoodsImport'
                          ELSE
                              'SantoGoodsImport'
                          
                          END AS tradeLocation
                          FROM importHistory
                          WHERE Procedure = '4000' or Procedure = '4071' or Procedure = '7100'")

dbWriteTable(mydb, "portsImport", portsImport, overwrite=TRUE)

portsImport_new <- dbGetQuery(mydb, "SELECT tradeLocation, Year, sum(importValue) as importValue FROM portsImport GROUP BY tradeLocation, Year ORDER BY Year ")

portsImport_new_pivot <- portsImport_new %>%
                         pivot_wider(names_from = tradeLocation, values_from = importValue)

#Merge trade imports and exports
tradeInGoods <- merge(portsExport_new_pivot, portsImport_new_pivot, by = 'Year')

#Calculate trade import total and export total
tradeInGoods$goodsImportTotal <- tradeInGoods$VilaGoodsImport + tradeInGoods$SantoGoodsImport
tradeInGoods$goodsExportTotal <- tradeInGoods$VilaGoodsExport + tradeInGoods$SantoGoodsExport

#Read Trade in services from the SQLite database 
tradeInServices <- dbGetQuery(mydb, "SELECT * FROM tblservices")

#Merge trade in goods and trade in services 
tradeGoodsAndServices <- merge(tradeInGoods, tradeInServices, by = "Year") 

tradeGoodsAndServices$overallExport <- tradeGoodsAndServices$goodsExportTotal + tradeGoodsAndServices$exportServices
tradeGoodsAndServices$overallImport <- tradeGoodsAndServices$goodsImportTotal + tradeGoodsAndServices$importServices
tradeGoodsAndServices$tradeBalance <- tradeGoodsAndServices$overallExport - tradeGoodsAndServices$overallImport

dbWriteTable(mydb, "tradeGoodsAndServices", tradeGoodsAndServices, overwrite=TRUE)

```


```{r Trade in goods by Port}
tradeGoodsAndServices %>%
  group_by(Year) %>%
  summarise(GExport_Lug = sum(SantoGoodsExport),
            GImport_Lug = sum(SantoGoodsImport),
            GExport_Vil = sum(VilaGoodsExport),
            GImport_Vil = sum(VilaGoodsImport),
            GVila_Balance = sum(VilaGoodsExport - VilaGoodsImport),
            GLug_Balance = sum(SantoGoodsExport - SantoGoodsImport), .groups="keep") %>%
            #POrts_Balance = sum(Balance), .groups="keep") %>%
  ungroup() %>%
   
  plot_ly(x = ~Year,
          y = ~GExport_Lug,
          color = "blue",
          type = "scatter", 
          mode = "lines+markers",
          name = "Luganville Goods Export") %>%
  
  add_trace(x = ~Year,
            y = ~GImport_Lug,
            color = "red",
            name = "Luganville Goods Import") %>%
  
  add_trace(x = ~Year,
            y = ~GExport_Vil,
            color = "green",
            type = "scatter",
            mode = "lines+markers",
            name = "Port Vila Goods Export") %>%
  
   add_trace(x = ~Year,
            y = ~GImport_Vil,
            color = "purple",
            type = "scatter",
            mode = "lines+markers",
            name = "Port Vila Goods Import") %>%
  
  plotly::layout(xaxis = list(title = "Year"), yaxis = list(title = 'Value (Billion (vatu)'))

```

### Trade Balance in Goods by Port

```{r Trade in goods Balances by Port}
tradeGoodsAndServices %>%
  group_by(Year) %>%
  summarise(GLug_Bal = sum(SantoGoodsExport - SantoGoodsImport),
            GVil_Bal = sum(VilaGoodsExport - VilaGoodsImport), .groups = "keep") %>%
  ungroup() %>%
   
  plot_ly(x = ~Year,
          y = ~GLug_Bal,
          color = "blue",
          type = "scatter", 
          mode = "lines+markers",
          name = "Luganville Trade Balance") %>%
  
  add_trace(x = ~Year,
            y = ~GVil_Bal,
            color = "red",
            name = "Port Vila Trade Balance") %>%
  
  
  plotly::layout(xaxis = list(title = "Year"), yaxis = list(title = 'Value (Billion (vatu)'))

```

Column {data-width=350}
-----------------------------------------------------------------------

### Trade Import Activities with 15 top countries overtime

```{r animated graphic for top 10 countries by year for imports, eval = TRUE}

#Extract export data from exportHistory from SQLite database
countryImport <- dbGetQuery(mydb, "SELECT importHistory.Year,                                                                                importHistory.CTY_Origin,
                                          country.CNTY_DESCR,
                                          sum(importHistory.Value)/1000000 as importValue
                                   FROM importHistory
                                   INNER JOIN country ON importHistory.CTY_Origin = country.CNTY_CODE
                                   WHERE CTY_Origin <> 'VR'
                                   GROUP BY Year, CNTY_DESCR
                                   ORDER BY Year ")


#imp <- dbGetQuery(mydb, "SELECT * FROM country_year_import ORDER BY Year")

countryImport %>% 
  mutate_if(is.factor, as.character) %>%
  mutate_if(is.integer, as.numeric) -> countryImport 

imp_formatted <- countryImport %>%
  group_by(Year) %>%
  mutate(rank = rank(-importValue),
         Value_rel = importValue/importValue[rank==1],
         Value_lbl = paste0(" ",round(importValue))) %>%
  group_by(CNTY_DESCR) %>% 
  filter(rank <=15) %>%
  ungroup()

anim_imp <- ggplot(imp_formatted, 
                   aes(rank, group = CNTY_DESCR, fill = as.factor(CNTY_DESCR), color = as.factor(CNTY_DESCR))) +
    
  geom_tile(aes(y = importValue/2, height = importValue, width = 0.9),
            alpha = 0.8, color = NA) +
  
  geom_text(aes(y = 0, label = paste(CNTY_DESCR, " ")), vjust = 0.2, hjust = 1) +
  geom_text(aes(y=importValue,label = Value_lbl, hjust=0)) +
  
  coord_flip(clip = "off", expand = FALSE) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_reverse() +
  guides(color = FALSE, fill = FALSE) +
  
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position="none",
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x = element_line( size=.1, color="grey" ),
        panel.grid.minor.x = element_line( size=.1, color="grey" ),
        plot.title=element_text(size=25, hjust=0.5, face="bold", colour="grey", vjust=-1),
        plot.subtitle=element_text(size=18, hjust=0.5, face="italic", color="grey"),
        plot.caption =element_text(size=8, hjust=0.5, face="italic", color="grey"),
        plot.background=element_blank(),
        plot.margin = margin(2,2, 2, 4, "cm")) +
  
  transition_states(Year, transition_length = 2, state_length = 1, wrap = FALSE) +
  
  view_follow(fixed_x = TRUE)  +
  
  labs(title = 'Import per Year : {closest_state}',  
       subtitle  =  "Top 15 Countries Vanuatu has imported from",
       caption  = "Import Value in Millions of Vatu | Data Source: Asycuda")

animate(anim_imp, nframes = 100, fps = 5)
```

### Trade Export Activities with 15 top countries overtime

```{r animated graphic for top 10 countries by year for exports, eval = TRUE}

#Extract import data from exportHistory from SQLite database
countryExport <- dbGetQuery(mydb, "SELECT exportHistory.Year,                                                                                exportHistory.CTY_Dest,
                                          country.CNTY_DESCR,
                                          sum(exportHistory.Value)/1000000 as exportValue
                                   FROM exportHistory
                                   INNER JOIN country ON exportHistory.CTY_Dest = country.CNTY_CODE
                                   WHERE CTY_Dest <> 'VR'
                                   GROUP BY Year, CNTY_DESCR
                                   ORDER BY Year ")

#Create import annimated column graph
countryExport %>%
  mutate_if(is.factor, as.character)%>%
  mutate_if(is.integer, as.numeric) -> countryExport 

exp_formatted <- countryExport %>%
  group_by(Year) %>%
  mutate(rank = rank(-exportValue),
         Value_rel = exportValue/exportValue[rank==1],
         Value_lbl = paste0(" ",round(exportValue))) %>%
  group_by(CNTY_DESCR) %>% 
  filter(rank <=15) %>%
  ungroup()

anim_exp <- ggplot(exp_formatted, 
                   aes(rank, group = CNTY_DESCR, fill = as.factor(CNTY_DESCR), color = as.factor(CNTY_DESCR))) +
  
  geom_tile(aes(y = exportValue/2, height = exportValue, width = 0.9),
            alpha = 0.8, color = NA) +
  
  geom_text(aes(y = 0, label = paste(CNTY_DESCR, " ")), vjust = 0.2, hjust = 1) +
  geom_text(aes(y=exportValue,label = Value_lbl, hjust=0)) +
  
  coord_flip(clip = "off", expand = FALSE) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_reverse() +
  guides(color = FALSE, fill = FALSE) +
  
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position="none",
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major.x = element_line( size=.1, color="grey" ),
        panel.grid.minor.x = element_line( size=.1, color="grey" ),
        plot.title=element_text(size=25, hjust=0.5, face="bold", colour="grey", vjust=-1),
        plot.subtitle=element_text(size=18, hjust=0.5, face="italic", color="grey"),
        plot.caption =element_text(size=8, hjust=0.5, face="italic", color="grey"),
        plot.background=element_blank(),
        plot.margin = margin(2,2, 2, 4, "cm")) +
  
  transition_states(Year, transition_length = 4, state_length = 1, wrap = FALSE) +
  
  view_follow(fixed_x = TRUE)  +
  
  labs(title = 'Export Per Year : {closest_state}',  
       subtitle  =  "Top 15 Countries Vanuatu has exported to",
       caption  = "Export Value in Millions of Vatu | Data Source: Asycuda") 

animate(anim_exp, nframes = 100, fps = 5)
```


Export by Commodity {data-navmenu="Trade in Export"}
=====================================

```{r Export by goods of interest}

historyExport <- dbGetQuery(mydb, "SELECT * FROM historyExport")
majorTen_Extract_Class <- dbGetQuery(mydb, "SELECT HS, [Principle.Exports] FROM ImpoprtExport_principal")  

majorTen_Exp <- majorTen_Extract_Class

majorTen_Exp_Select <- majorTen_Exp %>%
  distinct(Principle.Exports) %>%
  pull(Principle.Exports)

historyExport_Major10_merge <- merge(historyExport, majorTen_Exp, by="HS")

```


Column {.sidebar}
------------------------------------------------------------------
```{r}
selectInput(inputId = "MJTN_Exp",label = "Major 10 Goods",choices = majorTen_Exp_Select)

```

```{r}

updateData_export_commodity <- reactive(
  historyExport_Major10_merge %>% 
    filter(Principle.Exports == input$MJTN_Exp) %>%
    group_by(Year) %>%
    summarise(sum = sum(myvalue/1000000), .groups = "drop")
  )

```

column
---------------------------------------------------------------
### Bar Plot of Export by Commodity

```{r}
renderPlot({
  updateData_export_commodity() %>% 
  ggplot(aes(x=Year, y=sum)) +
    geom_col() +
    theme(panel.background = element_rect(fill = '#cccccc', colour = 'red')) +
    xlab("Year") +
    ylab("Value (Billion(Vatu))")
})
```


column
-------------------------------------------------
### Table display of the selected commodity 
```{r}
renderDT(
  updateData_export_commodity(), rownames = FALSE, extensions = 'Buttons', filter="top", editable=TRUE,
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    lengthMenu = list(c(10,50,100,-1),c(10,50,100,"All"))
  )
)

```

Export by Country {data-navmenu="Trade in Export"}
=====================================  

Column {.sidebar}
------------------------------------------------------------------
```{r}

uniqueCountriesISO2 <- unique(historyExport$CTY_Dest[is.na(historyExport$CTY_Dest) == FALSE])
countries <- countrycode(uniqueCountriesISO2, origin = "iso2c", destination = "country.name")

countriesToSelectFrom <- uniqueCountriesISO2
names(countriesToSelectFrom) <- countries
countriesToSelectFrom <- countriesToSelectFrom[is.na(countries) == FALSE]

selectInput(inputId = "exportCountry",label = "Metric", choices = countriesToSelectFrom)

```


```{r}

updateData_export_country <- reactive(
  historyExport %>% 
    filter(CTY_Dest == input$exportCountry) %>%
    group_by(Year) %>%
    summarise(sum = myvalue/1000000, .groups = "drop")
)

```

column
-------------------------------------------------
### Bar Plot of Export by Country

```{r}
renderPlot({
 updateData_export_country() %>%
  ggplot(aes(x=Year, y=sum)) +
  geom_col() +
  theme(panel.background = element_rect(fill = '#cccccc', colour = 'red')) +
  xlab("Year") +
  ylab("Value (Billion(Vatu))")
})
```


column
-------------------------------------------------
### Table display of the selected country 
```{r}
renderDT(
  updateData_export_country(), rownames = F, extensions = 'Buttons', filter="top", editable=T,
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    lengthMenu = list(c(10,50,100,-1),c(10,50,100,"All"))
  )
)

```


Import by Goods {data-navmenu="Trade in Import"}
=====================================

```{r Import by goods of interest}

majorTen_Import_Class <- dbGetQuery(mydb, "SELECT HS, [Principle.Imports] FROM ImpoprtExport_principal")
historyImport <- dbGetQuery(mydb, "SELECT * FROM historyImport")

historyImport_Major10_merge <- merge(historyImport, majorTen_Import_Class, by="HS")

majorTen_Imp <- majorTen_Import_Class %>%
  select(HS, Principle.Imports)

majorTen_Imp_Select <- majorTen_Imp %>%
  group_by(Principle.Imports) %>%
  summarise()

```

Column {.sidebar}
------------------------------------------------------------------
```{r}
selectInput(inputId = "MJTN",label = "Major 10 Goods",choices = majorTen_Imp_Select)

```

```{r}

updateData_import_commodity <- reactive(
  historyImport_Major10_merge %>% 
    filter(Principle.Imports == input$MJTN) %>%
    group_by(Year) %>%
    summarise(sum = sum(myvalue/1000000), .groups = "drop")
  )

```

column
-------------------------------------------------
### Bar Plot of Import by Commodity

```{r}
renderPlot({
  updateData_import_commodity() %>% 
  ggplot(aes(x=Year, y=sum)) +
    geom_col() +
    theme(panel.background = element_rect(fill = '#cccccc', colour = 'red')) +
    xlab("Year") +
    ylab("Value (Billion(Vatu))")
})
```


column
-------------------------------------------------
### Table display of the selected commodity 
```{r}
renderDT(
  updateData_import_commodity(), rownames = F, extensions = 'Buttons', filter="top", editable=T,
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    lengthMenu = list(c(10,50,100,-1),c(10,50,100,"All"))
  )
)

```

Import by Country {data-navmenu="Trade in Import"}
===================================== 

Column {.sidebar}
------------------------------------------------------------------
```{r}

uniqueCountriesISO2 <- unique(historyImport$CTY_Origin[is.na(historyImport$CTY_Origin) == FALSE])
countries <- countrycode(uniqueCountriesISO2, origin = "iso2c", destination = "country.name")

countriesToSelectFrom <- uniqueCountriesISO2
names(countriesToSelectFrom) <- countries
countriesToSelectFrom <- countriesToSelectFrom[is.na(countries) == FALSE]

selectInput(inputId = "importCountry",label = "Metric", choices = countriesToSelectFrom)
```


```{r}

updateData_import_country <- reactive(
  historyImport %>% 
    filter(CTY_Origin == input$importCountry) %>%
    group_by(Year) %>%
    summarise(sum = myvalue/1000000, .groups = "drop") 
    
)

```


column
-------------------------------------------------
### Bar Plot of Import by Country

```{r}
renderPlot({
 updateData_import_country() %>%
  ggplot(aes(x=Year, y=sum)) +
  geom_col() +
  theme(panel.background = element_rect(fill = '#cccccc', colour = 'red')) +
  xlab("Year") +
  ylab("Value (Billion(Vatu))")
})
```

column
-------------------------------------------------
### Table display of the selected country 
```{r}
renderDT(
  updateData_import_country(), rownames = F, extensions = 'Buttons', filter="top", editable=T,
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    lengthMenu = list(c(10,50,100,-1),c(10,50,100,"All"))
  )
)

```

Trade through Maps
===============================================



```{r load world map, include = FALSE, eval= FALSE}
# Download world country boundaries from: https://www.naturalearthdata.com/downloads/50m-cultural-vectors/50m-admin-0-countries-2/ (https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/50m/cultural/ne_50m_admin_0_countries.zip)
# Get a map of the world as as a spatial polygons object
shapeFile <- file.path("ne_50m_admin_0_countries", "ne_50m_admin_0_countries.shp")
world <- readOGR(shapeFile, verbose = FALSE)

# Simplify the polygons
worldSimple <- ms_simplify(world, keep = 0.1)

# Keep necessary columns for joining
worldSimple@data <- worldSimple@data[, c("NAME", "CONTINENT", "ISO_A2", "ISO_A3")]
```

```{r join world with trades data}
# Join in the imports data
worldImports <- merge(worldSimple, imp[imp$Year == 2018, c("CTY_ID", "Value_mill")], by.x = "ISO_A2", by.y = "CTY_ID")

# Join in the exports
worldImportsAndExports <- merge(worldImports, exp[exp$Year == 2018, c("CNTY_CODE", "Value_mill")],
                                by.x = "ISO_A2", by.y = "CNTY_CODE",
                                suffixes = c("_Imports", "_Exports"))
```

```{r build interactive map}
# Create a colour palette
valueRange <- range(c(worldImportsAndExports$Value_mill_Imports, 
                      worldImportsAndExports$Value_mill_Exports), na.rm = TRUE)
colours <- c("blue", "white", "red")
colourPalette <- colorNumeric(colours, valueRange)
colourPaletteWithoutNA <- colorNumeric(colours, valueRange, na.color=rgb(0,0,0,0))

# Create labels that will appear on hover
labels <- paste0("<strong>", worldImportsAndExports$NAME, "</strong>",
                 "<br>Import value: ", round(worldImportsAndExports$Value_mill_Imports, digits=2),
                    ifelse(is.na(worldImportsAndExports$Value_mill_Imports), "", " million"),
                 "<br>Export value: ", round(worldImportsAndExports$Value_mill_Exports, digits=2),
                 ifelse(is.na(worldImportsAndExports$Value_mill_Exports), "", " million")) %>% lapply(htmltools::HTML)

# Create an interactive map
leaflet(worldImportsAndExports, options = leafletOptions(crs = leafletCRS(crsClass = "L.CRS.EPSG4326"))) %>%

  setView(lng = 0, lat = 0, zoom = 1) %>%
  
  addPolygons(weight = 2, color = "white", dashArray = "3", fillOpacity = 0.5,
              fillColor = ~colourPalette(Value_mill_Imports),
              highlight = highlightOptions(weight = 5, color = "#666", dashArray = "",
                                           fillOpacity = 1, bringToFront = TRUE),
              label = labels,
              labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"),
                                          textsize = "15px", direction = "auto"),
              group = "Imports") %>%

  addPolygons(weight = 2, color = "white", dashArray = "3", fillOpacity = 0.5,
              fillColor = ~colourPalette(Value_mill_Exports),
              highlight = highlightOptions(weight = 5, color = "#666", dashArray = "",
                                           fillOpacity = 1, bringToFront = TRUE),
              label = labels,
              labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"),
                                          textsize = "15px", direction = "auto"),
              group = "Exports") %>%
  
  addLegend(values = valueRange, bins = seq(from = 0, to = signif(valueRange[2], 2), length.out = 5),
            pal = colourPaletteWithoutNA, title = "Value (millions)",
            position = "bottomleft", na.label = "") %>%
  
  addLayersControl(baseGroups = c("Imports", "Exports"), 
                   options = layersControlOptions(collapsed=FALSE)) %>%
  hideGroup(c("Exports"))
```








Export Microdata {data-navmenu="Microdata Download"}
==================================================================






Import Microdata {data-navmenu="Microdata Download"}
==================================================================

